CODECHECK: an Open Science initiative for the independent execution of computations underlying research articles during peer review to improve reproducibility

### 2021-05-27 reviews/response-round1.md

<https://doi.org/10.5256/f1000research.54932.r82470>

We thank the reviewers for their comments on our manuscript.  The
comments are below.  Indented are our responses.

## Reviewer 1

In this article, authors propose to implement a procedure to check for
the code accompanying a submission to a journal. To do so, they
describe a pipeline made of 6 steps that ultimately lead to the
delivery of a code check certificate meaning that someone external to
the author's lab has managed to re-run the code. At this point, no
checking that the results are correct is necessary. The authors
already issued several codecheck certificates in different
disciplines. I find the idea really nice and certainly necessary but
I've a few questions (even though some of them are already addressed
in the "limitations" section). Given the structure of the paper, I'll
just list my questions here:

1. How does CODECHECK compare to ACM Artifact reviews badges?
(https://www.acm.org/publications/policies/artifact-review-and-badging-current)

    These badges, introduced in August 2020, show whether code is
    available (different levels) and reproduces same results.  In
    principle CODECHECKER could award these badges (artifacts
    evaluated, functional).  We have made a note to this effect in the
    manuscript (Related work, end of paragraph 1.)

2. What would be the incentive for someone to code check the code? Being
aware of the increasing difficulty in finding reviewers, I don't think
it would be easy to recruit people to perform a task that can rapidly
become very technical and time consuming.

    This was addressed in our section "Who's got time for more peer
    review?" We would however note that we have a pool of about 20
    volunteers currently willing to do codechecks.

3. How do you handle the case when specific hardware is necessary
(e.g. NVidia GPU)? Is it documented somewhere such that code-checkers
might first verify if they have the necessary hardware to run the
code?

    This was handled in the limitation "What about my proprietary
    software and sensitive data." but we now mention hardware too in
    the first paragraph of that section.

4. How do you establish a check has failed? For example, what happens if
a code-checker gets a segfault (for some unknown reason) and the
author is unable to help. Is it deemed failed?

    We hope that codechecker and author can resolve problems, but in
    the end there may be problems that cannot be solved.  Open
    infrastructure could help as both author and codechecker can work
    together in the same environment to minimise these failures.
    Ultimately however, there may be failures, which are noted in the
    section "How are failures during checks handled?".

5. Who will pay for the computing resources needed to run heavy
simulations and/or to acquire necessary software such as e.g. Matlab?
When a simulation consumes a lot of resources, it might wise to give
the checker access to computing resources. This could be paid for by
the journal.

    In the section "Who will pay for compute time?" we mention this
	problem, and that toy examples might alleviate the need to re-run
	resource-intensive computations.  We agree that one model might be
	that a journal provide some resource for this service.  Likewise,
	in the following paragraph, we describe that our pragmatic
	approach for now is to find codecheckers that have access to
	particular software, e.g. MATLAB.

6. I did not see in the report example a description of the environment
necessary to run the software. How did you solve the "dependency
hell"? Since the code might break at some point in the future because
of incompatibility in some libraries or environments, it would be
necessary to have a mechanism describing the running environment such
that it can be re-run later.

    The short answer is "we didn't.".  In the paragraph "Should
    CODECHECK requirements be more demanding?" we note our low bar of
    simply getting a codecheck to run once.  We do, however, encourage
    CODECHECKERS to describe the environment in free text form in
    their report.  Moving towards machine-readable descriptions would
    be a natural extension.

7. What do you recommend if the reviews are both excellents but the code
check failed? Does this mean the paper is blocked until code check
passes or rejected or else?

   This is up to the editor of the journal -- see the "Importance"
   dimension of Figure 3.  At one end, it could indeed be a "strict
   requirement" to get a codecheck certificate for the paper to be
   accepted.  On the other hand, it could be entirely optional.

8. The code check proposal is close to some extents to the Journal of
Open Science Software where each reviewer is assigned a list of things
to check during the review. Do authors consider this pipeline when
establishing their own pipeline?

   We have not considered this pipeline, nor do we have an explicit
   idea.  We now note this reviewer list  at the end of the third paragraph
   of "Related Work".

9. To what extent the codecheck certificate can be updated automatically
via some kind of "manual continuous-integration"? I mean that when
reading a paper online, would it be possible to click a button to test
if the code still runs considering the latest versions of libraries?
(for example, the certificate has been issued for Python 2 but I want
to know if this is usable with Python 3).

    To follow on from point 6, this would make a natural extension, but for
    now we are still considering one point in time, and keeping the
    requirements as close to the authors as we can.

10. When you look at journals advertising open data policies, it is
unfortunately not rare to find articles in these same journals without
the actual data. Do you have some suggestion for educating editors to
actually enforce the code check a journal adopt it?

    We share this concern, and unfortunately have no simple
    suggestions for helping editors.  At this early stage, we think
    the approach should be one of encouraging uptake, rather than
    mandating it.  We also hope that having specific in-house
    experience, e.g. editorial staff to examine for code and data
    availability, can note this.  But at the end of the day, this
    again is dependent on the journal's workflow.

Some suggestions:

11. The badge that is delivered would need some time information since the
check is valid at one point in time (with a given software stack) and
does not guarantee future runs.

    Great idea. we could add the certificate number to the URL, or add
    the certificate number.  We will try to implement this when
    revising our workflows.

12. For specialized journals, you could consider to offer a common generic
environment where a code could be first tested. It this fails, then
you would need only to slightly modify the environment to add missing
dependencies. For example, in neuroscience, a Neuro Debian would
probably suit the needs of a large number of models.

    Yes. We will certainly bear this in mind in future work.

13. As editor-in-chief of ReScience C, I would like to inform authors
that the journal now accepts "reproduction report". The idea it to try
to re-run the code accompanying a published article and to report if
it succeeded or failed. Our own procedure to check for reproduction is
not standardized and we'll certainly benefit from the code check
initiative.

    Thank you for noting this.  We now mention the reproduction report
    in the manuscript where we describe Rescience C.

Overall, it's nice to have a clean description of a pipeline to check
for code even though some questions need to be addressed. Also, I'm
not too confident that journals will adopt it immediately and I'm
afraid such initiative will take time to be generalized. But we have
to start somewhere.

    We share your realistic assessment that (a) journals may be slow
    to adopt but that (b) we should start somewhere.

## Reviewer 2

1. Is there any community feedback on the transparency and
reusability of research published with CODECHECK certificates?

This is perhaps too big of an ask this early in the
initiative as research reuse and citations are independent factors of
the publication and peer review of this specific paper - but I’d still
be interested in any insights the authors have to offer on this topic.

    This is an excellent question.  It is perhaps too early for us to
	assess this given most certificates are under a year old, but we
	are not aware of any reuse yet of the codecheck-deposited
	material.  However, we certainly think it interesting to try and
	monitor this over a longer (3-5 year) timescale if possible.   As
	well as looking for citations of certificates, we could also check
	for forks of our repositories, and download statistics from
	Zenodo.

### 2021-06-15 codecheck-paper.tex
revision after first round of reviews
Transferred from https://www.overleaf.com/project/609a7da53dfe13bf1a4b42b9

\usepackage{fontspec}
\usepackage{fontawesome}
\newcommand{\github}[1]{\href{https://github.com/#1}{\faGithub}}
\newcommand{\twitter}[1]{\href{https://twitter.com/#1}{\faTwitter}}

@@

\begin{document}
\title{CODECHECK: An Open Science initiative for the independent
  execution of computations underlying research articles during peer
  review to  improve reproducibility}
\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025} \github{sje30} \twitter{StephenEglen}}
\affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, UK; \email{sje30@cam.ac.uk}}
\affil[$\ast$]{Corresponding authors}
\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}.}
\maketitle
\begin{abstract}
%\textbf{Background:}
%\textbf{Methods:}
%\textbf{Results:}
%\textbf{Conclusions:}
  The traditional scientific paper falls short of effectively communicating computational research.
 To help improve this situation, we propose a system by which the computational workflows underlying research articles are checked.
  The CODECHECK system uses open infrastructure and tools and can be integrated into review and publication processes in multiple ways.
  We describe these integrations along multiple dimensions (importance, who, openness, when).
  In collaboration with academic publishers and conferences, we demonstrate CODECHECK with 25 reproductions of diverse scientific publications.
  These CODECHECKs show that asking for reproducible workflows during a collaborative review can effectively improve executability.
  While CODECHECK has clear limitations, it may represent a building block in Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts.
\end{abstract}

\subsection*{Keywords}

reproducible research, Open Science, peer review, reproducibility, code sharing, data sharing, quality control, scholarly publishing

@@

Given that data are not routinely shared, what hope might there be for
sharing computer programs?  Both data and software are required to
validate a computational analysis; data can be seen as inert whereas
code requires an environment to be run in.  This makes software harder
to share.  Our experience is that researchers offer several reasons
for why code is not shared, e.g.,~``there is no documentation'', ``I
cannot maintain it'', or ``I do not want to give away my code to
competitors''.  Our view is that sharing code, wherever possible, is
good for the community and the individual
\cite{Barnes2010-iv,markowetz_five_2015}.  Having code and data openly
available, and archived, provides a valuable resource for others to
learn from, even if the code is broken or lacks documentation.
However, with a little effort, we believe that if an independent
person can re-run the programs, this is worth documenting and that
this reduces the barrier to evaluating non-text research materials.
Just as data journals' validations of data and all journals' peer
review provides a ``baseline reassurance'', i.e.,~that a paper has
been checked by someone with an understanding of the topic
\cite{fyfe_mission_2019}, the same baseline could be provided for the
workflow underlying a paper.  With this in mind, we have developed a
set of principles and an example workflow that provides a pragmatic
way of checking that a paper's code works, i.e., it is reproducible
following the Claerbout/Donoho/Peng terminology
\cite{barba_terminologies_2018}.

@@

\subsection*{Workflow and people}\label{workflow-people}

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people:
(1) the \textbf{author} of a paper providing the code to be checked,
(2) the \textbf{publisher} of a journal interested in publishing the author's paper, and
(3) the \textbf{codechecker}, who checks that the author's code works.
The six-step workflow we have refined is shown in
Figure~\ref{fig:workflow}.  In this article, we also refer to a
\textbf{peer-reviewer} who is independent of this process, and
performs the traditional academic review of the content of an article.

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1tGOZFNZle-oE1Ynw_tLZFmNcv3wgxD0HPs5bXzDOrng/edit
% emojis: https://github.com/googlefonts/noto-emoji and https://github.com/twitter/twemoji
% emoji codes:
% book stack: 1f4da
% package: 1f4e6
% file drawer: 1f5c4
% disk: 1f4be
% document: 1f4c4
% phd: 1f393
% detective: u1f575
% laptop: 1f4bb
% folder: 1f4c1 1f4c2
% receipt: 1f9fe
  \centering
      \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
      \caption{The CODECHECK example process implementation.
        Codecheckers act as detectives: They investigate and record,
        but do not fix issues.  Numbers in bold refer to steps
        outlined in the text.}
  \label{fig:workflow}
\end{figure}

\textbf{Step 1:} The author submits their manuscript along with the
code and data to the publisher.  The code and data need not be openly
available at this point.  However, in many cases the code and data may
be published on a code hosting platform, such as GitHub or
GitLab. Ideally, the author is expecting the CODECHECK and prepares
for it, e.g.,~by asking a colleague to attempt a reproduction, and
providing a set of instructions on how to re-run the workflow.

@@

\textbf{Step 3:} The codechecker runs the code, based on instructions
provided by the author. They check if some or all of the results from
the paper can be reproduced. If there are any problems running the
code, the codechecker asks the author for help, updates, or further
documentation.  The burden to provide reproducible material lies with
the author.  The codechecker then tries to run the code again.  This
process iterates until either the codechecker is successful, or the
codechecker concludes the workflow is not reproducible.  As part of
this process, the codechecker could work entirely locally, relying on
their own computing resources, or in the cloud, e.g.,~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives,
some of which are more tailored to scientific publications while
others offer commercial options for, e.g.,~publishers
(cf. \cite{konkol_publishing_2020}).  A cloud-based infrastructure
allows for the codechecker and author to collaboratively improve the
code and enforces a complete definition of the computing environment;
but, unless secure infrastructure is provided, e.g.,~by the publisher,
this requires the code and data to be published openly online.  Note
that the task of the codechecker is to check only the ``mechanics'' of the
workflow.  In the context of
mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish
between \emph{verification} and \emph{validation}; following their
definition, a CODECHECK ensures verification
%\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
of computational results, i.e.,~checking that code generates the
output it claims to create, but not a validation, i.e.,~checking that
the code implements the right algorithm to solve the specific research
problem.  Nevertheless, simply attempting to reproduce an output may
highlight a submission's shortcomings in meeting a journal's
requirements (cf.~\cite{christian_journal_2020}) and may effectively
increase transparency, thereby improving practices
(cf.~\cite{nosek_scientific_2012}) even if the check does not go into
every detail.

\textbf{Step 4:} The codechecker writes a certificate stating how the
code was run and includes a copy of outputs (figures or tables) that
were independently generated.  The certificate may include
recommendations on how to improve the material.  The free text in the
certificate can describe exactly what was checked, because each
workflow is unique.  Since no specific tool or platform is required,
such that no authors are excluded, it is futile for the
codechecker to use automation or fixed checklists.

@@

\textbf{Step 6:} The publisher can, depending on the timing, provide
the certificate to peer-reviewers or editors or publish it and link
between certificate, paper, and any repositories. Currently, the
codechecker creates these connections on Zenodo. They appear as links
with a relationship type on the Zenodo landing page for a certificate,
e.g., the ``related identifiers'' and ``alternate identifiers'' of
certificate \texttt{2020-025} \cite{cert-2020-025}.  The publisher
also credits the codechecker's work by depositing the activity
in scholarly profiles, such as ORCID (see peer review contributions in
ORCID records,
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}).
The publisher also ensures proper publication metadata, e.g.,~links
from the certificate repository to the published paper or the original
code repository.

@@

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1q2EdW3Ad-IpcD-CeoR1eIFMJW2fF6OVmDhCtCUmLO1k/edit
  \centering
      \includegraphics[width=0.8\textwidth]{figs/codecheck_dimensions.pdf}
  \caption{The dimensions of implementing a CODECHECK process.}
  \label{fig:dimensions}
\end{figure}

@@

The time at which a CODECHECK is done and its ascribed importance are
closely connected, so we describe the dimensions \emph{When} and
\emph{Importance} together.  The earlier a CODECHECK happens in the
publishing process, the more it can affect editorial decisions: Is a
paper published, sent back for revisions, or rejected?  Even earlier
checks, i.e.,~a CODECHECK of a preprint, may help to improve the
workflow itself, even before a publisher is involved.  As such,
codechecking papers could be part of a preprint server's policy or
initiated by interested authors.

@@

A \textbf{post-acceptance} CODECHECK would have the smallest impact on
editorial decisions and may simply provide \textbf{extra merit} on top
of the submission's acceptance.  This is the least impactful solution
in which all material is still evaluated and the results of the check
are properly acknowledged, because the check can be completed before
publication of the paper.  The GIScience checks (see
below) falls into this category: By displaying a badge on the volume
and article landing pages, the AGILE conference highlights articles
whose reproducibility was confirmed.  Similarly, in
collaborations with journals, some GIScience articles were
checked whilst authors worked on revisions.

@@

%The timing of a CODECHECK needs connecting to typical scholarly communication process.
%While ideas exist for radically changing how researchers can share their work
% \footnote{For example Octopus, \url{https://science-octopus.org/about}, or Hypergraph, \url{https://www.libscie.org/hypergraph}.},
% % mimosa is another platform idea, but no real prototype AFAICT:
% % https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
% % https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
% an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
Enhancing existing processes with CODECHECKs allows communities to
gradually transition towards more open practices.
When integrating a CODECHECK into existing review and publication processes, the \emph{turnaround time} is crucial.
Depending on when and who conducts the check, it might be done quickly
or it might delay publication.
We found that a CODECHECK generally takes  2--5 hours, with some outliers on the higher end.
This time includes writing and publishing the certificate but excludes
actual computation time, some of which took days.
These efforts are comparable to the time needed to peer review a submission,
which aligns with the efforts some volunteer codecheckers are willing to
make.
Currently, there is considerable amount of communicating about the
process, especially regarding who publishes which document when, so that proper
cross-referencing between paper and certificate is ensured via persistent
identifiers.  When integrated into a peer review platform, this
handling of documents should become much more streamlined.

@@

The technical nature of the check and the challenge of providing sufficient documentation is why we see great benefits in bidirectional communication between author and codechecker.
Instead of trying to
fix problems or guess the next step, the codechecker can ask the author to 
rework the documentation or update code.
Instead of struggling to provide perfect instructions and as a result possibly not sharing any code or data, the author can make a best effort to document sufficiently.
Authors and readers can profit from a codecheckers' experience and approach, as during the check they may create useful and instructive files, e.g.,~a machine-readable computing environment specification.
While communication between author and codechecker may be anonymised via
the publisher, it most likely only helps to protect the identity of
the codechecker, because code is hard to anonymise.
Therefore, the most effective and desirable situation for the
stakeholders is to hold a open and collaborative CODECHECK.
The contributions by the codechecker may even be integrated into
the code of the workflow and be acknowledged as code commits. This way, 
proper credit can be given within the research software development community.

@@

Just as with peer-reviewers, a potential codechecker should have the
right skills and availability to do the work.  Ideally, the
codechecker has a matching code \emph{and} domain expertise to the
paper, although a well-documented workflow should be executable by any
computationally-competent person. Naturally, the more
prerequisite knowledge the codechecker has, the quicker they can
understand the goals and mechanics of an analysis.  From our
experiences, the priority should be given to matching technical
expertise first, as lacking knowledge in setting up a computing
environment with a particular language or tool is much more of a
problem than assessing the outcome, e.g.,~comparing created figures
with the original, without an in-depth understanding of the domain.
The depth of the check will mostly be driven by the time required and
expertise of the checker, though in general, we expect a CODECHECK 
to consider reproducibility of the results above performance of the code.

Codecheckers could be drawn from a regular pool of
\textbf{peer-reviewers}, or from a special group of
\textbf{reproducibility reviewers} via specific roles such as
\textbf{reproducibility editors}, or \textbf{editorial staff} with a
publisher.  One codechecker is sufficient to verify the workflow since
it is mostly a factual process.  Code usually harbours systematic and
repeatable mistakes and is thereby more reliable and auditable than
processes controlled by humans \cite{tibav:42484}, e.g.,~in a
laboratory.  If however publication of the paper depends on the
CODECHECK, a second opinion may be required.

@@

The codechecker could be a member of \textbf{editorial staff}; this is
the most controlled but also resource-intensive option.  Such a
resource commitment would show that publishers are investing in
reproducibility, yet this commitment may be hard for small publishers.
These codecheckers could be fully integrated into internal workflows.
Credit for doing the codecheck is also achieved, as it is part of
their duties.  By contrast, it is useful for researchers to be
publicly credited for their reviewing activity.  A regular review
may be listed in public databases (e.g.,~ORCID, see Step~6 above,
or commercial offerings such as Publons, \url{https://publons.com/}, and
ReviewerCredits, \url{https://www.reviewercredits.com/});
a codechecker could be similarly listed.
% The number of scientific reviews is of course only a coarse indicator of an individual's community contribution, yet it is unclear if a regular reviewer, who conducts a regular scientific review and in addition also conducts a CODECHECK, should be credited twice.
% In this case, the published CODECHECK certificate and possibly the published peer review can clearly describe the possibly large efforts put into both.

The codechecker community has over 20 volunteers who signed up in the last
year, see \url{https://github.com/codecheckers/codecheckers/}.  Their
motivations, mentioned in the registration
information at \url{https://github.com/codecheckers/codecheckers/labels/registration},
include: supporting reproducible research and Open Science, improve
coding skills, gaining experience in helping scientists with their
code, encouraging a sharing culture, and learning from other people's
mistakes; many are also motivated simply by curiosity.  We see
benefits to an open shared list of codecheckers across journals rather
than a private in-house group, as this may allow for better matches
regarding expertise and workload sharing.  
This community can establish CODECHECK as a viable option
for independent no-cost Open Access journals.

>>>>

\newcommand{\rev}[1]{#1}
%\newcommand{\rev}[1]{\textit{#1}}
%\newcommand{\rev}[1]{\textcolor{blue}{#1}}

\usepackage{fontawesome}
\newcommand{\github}[1]{\href{https://github.com/#1}{\faGithub}}
\newcommand{\twitter}[1]{\href{https://twitter.com/#1}{\faTwitter}}

@@

\begin{document}
\title{CODECHECK: an Open Science initiative for the independent
  execution of computations underlying research articles during peer
  review to  improve reproducibility}
\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046}}
\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025}}
\affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, UK; \email{sje30@cam.ac.uk}}
\affil[$\ast$]{Corresponding authors}
\maketitle

\begin{abstract}
 The traditional scientific paper falls short of effectively communicating computational research.  To help improve this situation, we propose a system by which the computational workflows underlying research articles are checked. The CODECHECK system uses open infrastructure and tools and can be integrated into review and publication processes in multiple ways. We describe these integrations along multiple dimensions (importance, who, openness, when). In collaboration with academic publishers and conferences, we demonstrate CODECHECK with 25 reproductions of diverse scientific publications. These CODECHECKs show that asking for reproducible workflows during a collaborative review can effectively improve executability. While CODECHECK has clear limitations, it may represent a building block in Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts. The CODECHECK website can be accessed here: \url{https://codecheck.org.uk/}.
\end{abstract}

@@

Given that data are not routinely shared, what hope might there be for
sharing computer programs?  Both data and software are required to
validate a computational analysis; data can be seen as inert whereas
code requires an environment to be run in.  This makes software harder
to share.  Our experience is that researchers offer several reasons
for why code is not shared, e.g.,~``there is no documentation'', ``I
cannot maintain it'', or ``I do not want to give away my code to
competitors''.  Our view is that sharing code, wherever possible, is
good for the community and the individual
\cite{Barnes2010-iv,markowetz_five_2015}.  Having code and data openly
available, and archived, provides a valuable resource for others to
learn from, even if the code is broken or lacks documentation.
However, with a little effort, we believe that if an independent
person can re-run the programs, this is worth documenting and that
this reduces the barrier to evaluating non-text research materials.
Just as data journals' validations of data and all journals' peer
review provides a ``baseline reassurance'', i.e.,~that a paper has
been checked by someone with an understanding of the topic
\cite{fyfe_mission_2019}, the same baseline could be provided for the
\rev{computational} workflow underlying a paper.  With this in mind, we have developed a
set of principles and an example \rev{CODECHECK} workflow that provides a pragmatic
way of checking that a paper's code works, i.e., it is reproducible
following the Claerbout/Donoho/Peng terminology
\cite{barba_terminologies_2018}.

@@

\subsection*{\rev{CODECHECK w}orkflow and people}\label{workflow-people}

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The \rev{CODECHECK} workflow involves three
groups of people:
(1) the \textbf{author} of a paper providing the code to be checked,
(2) the \textbf{publisher} of a journal interested in publishing the author's paper, and
(3) the \textbf{codechecker}, who checks that the author's code works.
The six-step \rev{CODECHECK} workflow we have refined is shown in
Figure~\ref{fig:workflow}.  In this article, we also refer to a
\textbf{peer-reviewer} who is independent of this process, and
performs the traditional academic review of the content of an article.

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1tGOZFNZle-oE1Ynw_tLZFmNcv3wgxD0HPs5bXzDOrng/edit
% emojis: https://github.com/googlefonts/noto-emoji and https://github.com/twitter/twemoji
% emoji codes:
% book stack: 1f4da
% package: 1f4e6
% file drawer: 1f5c4
% disk: 1f4be
% document: 1f4c4
% phd: 1f393
% detective: u1f575
% laptop: 1f4bb
% folder: 1f4c1 1f4c2
% receipt: 1f9fe
  \centering
      \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
      \caption{The CODECHECK example \rev{workflow} implementation.
        Codecheckers act as detectives: They investigate and record,
        but do not fix issues.  Numbers in bold refer to steps
        outlined in the text.}
  \label{fig:workflow}
\end{figure}

\textbf{Step 1:} The author submits their manuscript along with the
code and data to the publisher.  The code and data need not be openly
available at this point.  However, in many cases the code and data may
be published on a code hosting platform, such as GitHub or
GitLab. Ideally, the author is expecting the CODECHECK and prepares
for it, e.g.,~by asking a colleague to attempt a reproduction, and
providing a set of instructions on how to re-run the \rev{computational} workflow.

@@

\textbf{Step 3:} The codechecker runs the code, based on instructions
provided by the author. They check if some or all of the results from
the paper can be reproduced. If there are any problems running the
code, the codechecker asks the author for help, updates, or further
documentation.  The burden to provide reproducible material lies with
the author.  The codechecker then tries to run the code again.  This
process iterates until either the codechecker is successful, or the
codechecker concludes the \rev{paper's} workflow is not reproducible.  As part of
this process, the codechecker could work entirely locally, relying on
their own computing resources, or in the cloud, e.g.,~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives,
some of which are more tailored to scientific publications while
others offer commercial options for, e.g.,~publishers
(cf. \cite{konkol_publishing_2020}).  A cloud-based infrastructure
allows for the codechecker and author to collaboratively improve the
code and enforces a complete definition of the computing environment;
but, unless secure infrastructure is provided, e.g.,~by the publisher,
this requires the code and data to be published openly online.  Note
that the task of the codechecker is to check only the ``mechanics'' of the
\rev{computational} workflow.  In the context of
mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish
between \emph{verification} and \emph{validation}; following their
definition, a CODECHECK ensures verification
%\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
of computational results, i.e.,~checking that code generates the
output it claims to create, but not a validation, i.e.,~checking that
the code implements the right algorithm to solve the specific research
problem.  Nevertheless, simply attempting to reproduce an output may
highlight a submission's shortcomings in meeting a journal's
requirements (cf.~\cite{christian_journal_2020}) and may effectively
increase transparency, thereby improving practices
(cf.~\cite{nosek_scientific_2012}) even if the check does not go into
every detail.

\textbf{Step 4:} The codechecker writes a certificate stating how the
code was run and includes a copy of outputs (figures or tables) that
were independently generated.  The certificate may include
recommendations on how to improve the material.  The free text in the
certificate can describe exactly what was checked, because each
\rev{computational} workflow is unique.  Since no specific tool or platform is required,
such that no authors are excluded, it is futile for the
codechecker to use automation or fixed checklists.

@@

\textbf{Step 6:} The publisher can, depending on the timing, provide
the certificate to peer-reviewers or editors or publish it and link
between certificate, paper, and any repositories. Currently, the
codechecker creates these connections on Zenodo. They appear as links
with a relationship type on the Zenodo landing page for a certificate,
e.g., the ``related identifiers'' and ``alternate identifiers'' of
certificate \texttt{2020-025} \cite{cert-2020-025}.  The publisher
also credits the codechecker's work by depositing the activity
in scholarly profiles, such as ORCID (see peer review contributions in
\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{ORCID records}).
The publisher also ensures proper publication metadata, e.g.,~links
from the certificate repository to the published paper or the original
code repository.

@@

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1q2EdW3Ad-IpcD-CeoR1eIFMJW2fF6OVmDhCtCUmLO1k/edit
  \centering
      \includegraphics[width=0.8\textwidth]{figs/codecheck_dimensions.pdf}
  \caption{The dimensions of implementing a CODECHECK \rev{workflow}.}
  \label{fig:dimensions}
\end{figure}

@@

The time at which a CODECHECK is done and its ascribed importance are
closely connected, so we describe the dimensions \emph{When} and
\emph{Importance} together.  The earlier a CODECHECK happens in the
publishing process, the more it can affect editorial decisions: Is a
paper published, sent back for revisions, or rejected?  Even earlier
checks, i.e.,~a CODECHECK of a preprint, may help to improve the
\rev{computational} workflow itself, even before a publisher is involved.  As such,
codechecking papers could be part of a preprint server's policy or
initiated by interested authors.

@@

A \textbf{post-acceptance} CODECHECK would have the smallest impact on
editorial decisions and may simply provide \textbf{extra merit} on top
of the submission's acceptance.  This is the least impactful solution
in which all material is still evaluated and the results of the check
are properly acknowledged, because the check can be completed before
publication of the paper.  The GIScience checks (see
below) falls into this category: by displaying a badge on the volume
and article landing pages, the AGILE conference highlights articles
whose reproducibility was confirmed.  Similarly, in
collaborations with journals, some GIScience articles were
checked whilst authors worked on revisions.

@@

%The timing of a CODECHECK needs connecting to typical scholarly communication process.
%While ideas exist for radically changing how researchers can share their work
% \footnote{For example Octopus, \url{https://science-octopus.org/about}, or Hypergraph, \url{https://www.libscie.org/hypergraph}.},
% % mimosa is another platform idea, but no real prototype AFAICT:
% % https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
% % https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
% an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
Enhancing existing \rev{review and publication} processes with CODECHECKs allows communities to
gradually transition towards more open practices.
When integrating a CODECHECK into existing review and publication processes, the \emph{turnaround time} is crucial.
Depending on when and who conducts the check, it might be done quickly
or it might delay publication.
We found that a CODECHECK generally takes  2--5 hours, with some outliers on the higher end.
This time includes writing and publishing the certificate but excludes
actual computation time, some of which took days.
These efforts are comparable to the time needed to peer review a submission,
which aligns with the efforts some volunteer codecheckers are willing to
make.
Currently, there is considerable amount of communicating about the \rev{CODECHECK workflow}, especially regarding who publishes which document when, so that proper
cross-referencing between paper and certificate is ensured via persistent
identifiers.  When integrated into a peer review platform, this
handling of documents should become much more streamlined.

@@

The technical nature of the check and the challenge of providing sufficient documentation is why we see great benefits in bidirectional communication between author and codechecker.
Instead of trying to
fix problems or guess the next step, the codechecker can ask the author to 
rework the documentation or update code.
Instead of struggling to provide perfect instructions and as a result possibly not sharing any code or data, the author can make a best effort to document sufficiently.
Authors and readers can profit from a codecheckers' experience and approach, as during the check they may create useful and instructive files, e.g.,~a machine-readable computing environment specification.
While communication between author and codechecker may be anonymised via
the publisher, it most likely only helps to protect the identity of
the codechecker, because code is hard to anonymise.
Therefore, the most effective and desirable situation for the
stakeholders is to hold a open and collaborative CODECHECK.
The contributions by the codechecker may even be integrated into
the code of the \rev{paper's} workflow and be acknowledged as code commits. This way, 
proper credit can be given within the research software development community.

@@

Just as with peer-reviewers, a potential codechecker should have the
right skills and availability to do the work.  Ideally, the
codechecker has a matching code \emph{and} domain expertise to the
paper, although a well-documented \rev{analysis} should be executable by any
computationally-competent person. Naturally, the more
prerequisite knowledge the codechecker has, the quicker they can
understand the goals and mechanics of an analysis.  From our
experiences, the priority should be given to matching technical
expertise first, as lacking knowledge in setting up a computing
environment with a particular language or tool is much more of a
problem than assessing the outcome, e.g.,~comparing created figures
with the original, without an in-depth understanding of the domain.
The depth of the check will mostly be driven by the time required and
expertise of the checker, though in general, we expect a CODECHECK 
to consider reproducibility of the results above performance of the code.

Codecheckers could be drawn from a regular pool of
\textbf{peer-reviewers}, or from a special group of
\textbf{reproducibility reviewers} via specific roles such as
\textbf{reproducibility editors}, or \textbf{editorial staff} with a
publisher.  One codechecker is sufficient to verify the \rev{paper's} workflow since
it is mostly a factual process.  Code usually harbours systematic and
repeatable mistakes and is thereby more reliable and auditable than
processes controlled by humans \cite{tibav:42484}, e.g.,~in a
laboratory.  If however publication of the paper depends on the
CODECHECK, a second opinion may be required.

@@

The codechecker could be a member of \textbf{editorial staff}; this is
the most controlled but also resource-intensive option.  Such a
resource commitment would show that publishers are investing in
reproducibility, yet this commitment may be hard for small publishers.
These codecheckers could be fully integrated into the internal \rev{publication process}.
Credit for doing the codecheck is also achieved, as it is part of
their duties.  By contrast, it is useful for researchers to be
publicly credited for their reviewing activity.  A regular review
may be listed in public databases (e.g.,~ORCID, see Step~6 above,
or commercial offerings such as \href{https://publons.com/about/home/}{Publons}, and
\href{https://www.reviewercredits.com/}{ReviewerCredits});
a codechecker could be similarly listed.
% The number of scientific reviews is of course only a coarse indicator of an individual's community contribution, yet it is unclear if a regular reviewer, who conducts a regular scientific review and in addition also conducts a CODECHECK, should be credited twice.
% In this case, the published CODECHECK certificate and possibly the published peer review can clearly describe the possibly large efforts put into both.

The codechecker community has over 20 volunteers who signed up in the last
year, see \url{https://github.com/codecheckers/codecheckers/}.  Their
motivations, mentioned in the \href{https://github.com/codecheckers/codecheckers/labels/registration}{registration
information},
include: supporting reproducible research and Open Science, improve
coding skills, gaining experience in helping scientists with their
code, encouraging a sharing culture, and learning from other people's
mistakes; many are also motivated simply by curiosity.  We see
benefits to an open shared list of codecheckers across journals rather
than a private in-house group, as this may allow for better matches
regarding expertise and workload sharing.  
This community can establish CODECHECK as a viable option
for independent no-cost Open Access journals.
